{\huge Supplements for: Changes and trends in budburst and leaf flush across Europe and North America} \\A meta-analysis of local adaptation in spring phenology studies

### 2023-10-04 Docs/supplement.Rnw
made small edits

\begin{document}
\SweaveOpts{concordance=TRUE} % For RStudio hiccups
\title{{\huge Supplements for: Changes and trends in budburst and leaf flush across Europe and North America} \\A meta-analysis of local adaptation in spring phenology studies}
\author{Ziyun Zeng \& E. M. Wolkovich}
\date{2023}
\maketitle

\section{Additional Methods}
\label{section:addmethods}
% @ alina_jan27: do we need to cite those paper indicated above that we didnt end up using through bibtex?? % EMW replies -- no, some journals will require a spreadsheet (what you discarded and why) though -- but until we pick a journal and know what's required we don't need to do any of that. 
We had to exclude several studies that reported spring events on a quantitative scale. This was because (1) such studies usually only assessed where on the scale the spring event of a tree fell onto on the same days across different years (e.g. Robson et al., 2013; Vander et al., 2015; Santini et al., 2014; Schueler \& Liesebach, 2004), and (2) scales are not always consistent across different studies (Chmura \& Rozkowski 2002; Dhont et al., 2010; Wang et al., 2022). Such factors made it impossible to convert the quantitative scale to DOY.
\\
\\
We additionally excluded several studies because we could not pinpoint their location, or because they focused on non-native species or elevational trends. We excluded studies that did not provide the exact latitude and longitude of the common garden (Bongarten, 1978) or the provenances (Hall et al., 2007; Soolanayakanahally et al., 2013), or they did not link the latitude and longitude of each provenance to the DOY of spring events (Deans \& Harvey, 1996). We also left out studies in which woody plants from North American provenances were planted in common gardens in Europe (Cannell et al., 1987; Lavadinovic et al., 2013) because we wanted to test continental variations. Finally, as we are focused on latitudinal trends we excluded studies that examined only provenance altitude (Vitasse et al., 2009; Vitasse et al., 2010; Li et al., 1997; Alberto et al., 2011; Acevedo-Rodríguez et al., 2006).

>>>>

\begin{document}
\SweaveOpts{concordance=TRUE} % For RStudio hiccups
\begin{flushright}
Version dated: \today
\end{flushright}

\bigskip
\medskip
\begin{center}

% Insert your title:
\noindent{\Large {\bf Supporting Information: Weak evidence of provenance effects in spring phenology across Europe and North America}}\\ 
\bigskip

\noindent {\normalsize \sc
Z. A. Zeng $^{1}$ \& E. M. Wolkovich$^{2}$}\\

\noindent {\small \it
$^1$ Forest Resources Management, Faculty of Forestry, University of British Columbia, 2424 Main Mall, Vancouver, BC V6T 1Z4\\
$^2$ Forest \& Conservation Sciences, Faculty of Forestry, University of British Columbia, 2424 Main Mall, Vancouver, BC V6T 1Z4}
\end{center}
\medskip
\noindent{\bf Corresponding author:} Z. A. Zeng, see $^{1}$ above ; E-mail: alinazengziyun@yahoo.com\\

% \noindent Additional information: Zeng (0009-0000-4647-9457); Wolkovich (0000-0001-7653-893X)\\

% \title{{\huge Supplements for: Changes and trends in budburst and leaf flush across Europe and North America} \\A meta-analysis of local adaptation in spring phenology studies}
% \author{Ziyun Zeng \& E. M. Wolkovich}
% \date{2023}
% \maketitle

\newpage
\section{Additional Methods}
\label{section:addmethods}
% @ alina_jan27: do we need to cite those paper indicated above that we didnt end up using through bibtex?? % EMW replies -- no, some journals will require a spreadsheet (what you discarded and why) though -- but until we pick a journal and know what's required we don't need to do any of that. 
We had to exclude several studies that reported spring events on a quantitative scale. This was because (1) such studies usually only assessed where on the scale the spring event of a tree fell onto on the same days across different years (e.g. Robson et al., 2013; Vander et al., 2015; Santini et al., 2014; Schueler \& Liesebach, 2004), and (2) scales are not always consistent across different studies (Chmura \& Rozkowski 2002; Dhont et al., 2010; Wang et al., 2022). Such factors made it impossible to convert the quantitative scale to DOY.
\\
\\
We additionally excluded several studies because we could not pinpoint their location, or because they focused on non-native species or elevational trends. We excluded studies that did not provide the exact latitude and longitude of the common garden (Bongarten, 1978) or the provenances (Hall et al., 2007; Soolanayakanahally et al., 2013), or they did not link the latitude and longitude of each provenance to the DOY of spring events (Deans \& Harvey, 1996). We also left out studies in which woody plants from North American provenances were planted in common gardens in Europe (Cannell et al., 1987; Lavadinovic et al., 2013) because we wanted to test continental variations. Finally, as we are focused on latitudinal trends we excluded studies that examined only provenance altitude (Vitasse et al., 2009; Vitasse et al., 2010; Li et al., 1997; Alberto et al., 2011; Acevedo-Rodríguez et al., 2006).

### 2023-10-04 Docs/supplement.tex
made small edits

\begin{document}
\input{/Users/alina/Documents/git/localadaptclim/Docs/figure_ms/demoFig-concordance} % For RStudio hiccups
\title{{\huge Supplements for: Changes and trends in budburst and leaf flush across Europe and North America} \\A meta-analysis of local adaptation in spring phenology studies}
\author{Ziyun Zeng \& E. M. Wolkovich}
\date{2023}
\maketitle

\section{Additional Methods}
\label{section:addmethods}
% @ alina_jan27: do we need to cite those paper indicated above that we didnt end up using through bibtex?? % EMW replies -- no, some journals will require a spreadsheet (what you discarded and why) though -- but until we pick a journal and know what's required we don't need to do any of that. 
We had to exclude several studies that reported spring events on a quantitative scale. This was because (1) such studies usually only assessed where on the scale the spring event of a tree fell onto on the same days across different years (e.g. Robson et al., 2013; Vander et al., 2015; Santini et al., 2014; Schueler \& Liesebach, 2004), and (2) scales are not always consistent across different studies (Chmura \& Rozkowski 2002; Dhont et al., 2010; Wang et al., 2022). Such factors made it impossible to convert the quantitative scale to DOY.
\\
\\
We additionally excluded several studies because we could not pinpoint their location, or because they focused on non-native species or elevational trends. We excluded studies that did not provide the exact latitude and longitude of the common garden (Bongarten, 1978) or the provenances (Hall et al., 2007; Soolanayakanahally et al., 2013), or they did not link the latitude and longitude of each provenance to the DOY of spring events (Deans \& Harvey, 1996). We also left out studies in which woody plants from North American provenances were planted in common gardens in Europe (Cannell et al., 1987; Lavadinovic et al., 2013) because we wanted to test continental variations. Finally, as we are focused on latitudinal trends we excluded studies that examined only provenance altitude (Vitasse et al., 2009; Vitasse et al., 2010; Li et al., 1997; Alberto et al., 2011; Acevedo-Rodríguez et al., 2006).

>>>>

\begin{document}
\input{/Users/alina/Documents/git/localadaptclim/Docs/figure_ms/demoFig-concordance} % For RStudio hiccups
\begin{flushright}
Version dated: \today
\end{flushright}

\bigskip
\medskip
\begin{center}

% Insert your title:
\noindent{\Large {\bf Supporting Information: Weak evidence of provenance effects in spring phenology across Europe and North America}}\\ 
\bigskip

\noindent {\normalsize \sc
Z. A. Zeng $^{1}$ \& E. M. Wolkovich$^{2}$}\\

\noindent {\small \it
$^1$ Forest Resources Management, Faculty of Forestry, University of British Columbia, 2424 Main Mall, Vancouver, BC V6T 1Z4\\
$^2$ Forest \& Conservation Sciences, Faculty of Forestry, University of British Columbia, 2424 Main Mall, Vancouver, BC V6T 1Z4}
\end{center}
\medskip
\noindent{\bf Corresponding author:} Z. A. Zeng, see $^{1}$ above ; E-mail: alinazengziyun@yahoo.com\\

% \noindent Additional information: Zeng (0009-0000-4647-9457); Wolkovich (0000-0001-7653-893X)\\

% \title{{\huge Supplements for: Changes and trends in budburst and leaf flush across Europe and North America} \\A meta-analysis of local adaptation in spring phenology studies}
% \author{Ziyun Zeng \& E. M. Wolkovich}
% \date{2023}
% \maketitle

\newpage
\section{Additional Methods}
\label{section:addmethods}
% @ alina_jan27: do we need to cite those paper indicated above that we didnt end up using through bibtex?? % EMW replies -- no, some journals will require a spreadsheet (what you discarded and why) though -- but until we pick a journal and know what's required we don't need to do any of that. 
We had to exclude several studies that reported spring events on a quantitative scale. This was because (1) such studies usually only assessed where on the scale the spring event of a tree fell onto on the same days across different years (e.g. Robson et al., 2013; Vander et al., 2015; Santini et al., 2014; Schueler \& Liesebach, 2004), and (2) scales are not always consistent across different studies (Chmura \& Rozkowski 2002; Dhont et al., 2010; Wang et al., 2022). Such factors made it impossible to convert the quantitative scale to DOY.
\\
\\
We additionally excluded several studies because we could not pinpoint their location, or because they focused on non-native species or elevational trends. We excluded studies that did not provide the exact latitude and longitude of the common garden (Bongarten, 1978) or the provenances (Hall et al., 2007; Soolanayakanahally et al., 2013), or they did not link the latitude and longitude of each provenance to the DOY of spring events (Deans \& Harvey, 1996). We also left out studies in which woody plants from North American provenances were planted in common gardens in Europe (Cannell et al., 1987; Lavadinovic et al., 2013) because we wanted to test continental variations. Finally, as we are focused on latitudinal trends we excluded studies that examined only provenance altitude (Vitasse et al., 2009; Vitasse et al., 2010; Li et al., 1997; Alberto et al., 2011; Acevedo-Rodríguez et al., 2006).
